---
title: "Food Web Comparisons"
output:
  pdf_document: default
  html_document: default
date: "2024-06-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(igraph) 
library(network) 
library(sna)
library(ggraph)
library(visNetwork)
library(threejs)
library(networkD3)
library(ndtv) 
library(tidyverse)
library(viridis)
library(cheddar)
library(kableExtra)
library(devtools)
#install_github("lsaravia/multiweb")
library(multiweb)
library(NetworkExtinction)
#install_github('SEELab/enaR', ref='develop') # using with caution since this package has been kicked off cran, but is still being actively?? developed on github
#library(enaR) #not working, oh well

#options(knitr.table.format = "latex")
```


```{r read in data}

nodes <- read_csv("data/foodweb_V8/trophic_nodes.csv")
links <- read_csv("data/foodweb_V8/trophic_links.csv")

```
# Time to compare foodwebs!
## Let's start with parasites vs no parasites

### First make subweb data
```{r make no parasite subwebs}
######################################
########### NO PARASITES #############
######################################

########### filter parasites out to create new node list ########### 
nodes_nopsites <- nodes %>% 
  filter(parasite == "n")

###########  subset the web using this new node list ########### 
# we only want consumers/resource links in the web if those consumers/resources are in our clean node list
# we can do this in five steps

#1. create a "keep" list of consumers
consumers_keep <- nodes_nopsites %>% 
  dplyr::select(node) %>%  # we only need the unique identifier, "node", for matching
  rename(consumer = node) # the unique identifier of consumers in the link list is "consumer", and these colnames must match

#2. create a version of the link list that only has links with consumers in our "keep" list, e.g. free living organisms only
consumer_keep_links <- inner_join(consumers_keep, links, by = join_by(consumer)) # inner join keeps only rows in 'links' that match 'consumers_keep'
  
#3. create a "keep" list of resources (this is identical to our consumer keep list, we just need to change the column name)
resources_keep <- consumers_keep %>% 
  rename(resource = consumer) # the unique identifier of resources in the link list is "resource", and these colnames must match

#4. use 'inner_join()' to create a link list from the consumer_keep_links list that ALSO only has resources from the keep list, e.g. free living organisms
links_nopsites <- inner_join(resources_keep, consumer_keep_links, by = join_by(resource))

```

### Then make cheddar and igraph objects
```{r network objects}

########### CHEDDAR ###########

## with parasites ##
# create properties list
properties.p <- list(title = "ca riz parasites") #syntax that the help doc suggests

# create community object
ca_riz_parasites <- Community(nodes = nodes, properties.p, trophic.links = links)
ca_riz_parasites

## no parasites ##

# create properties list
properties.np <- list(title = "ca riz no parasites") #syntax that the help doc suggests

# create community object
ca_riz_freeliving <- Community(nodes = nodes_nopsites, properties.np, trophic.links = links_nopsites)
ca_riz_freeliving

############ IGRAPH ############

## with parasites ##
igraph_links <- links %>% 
  relocate(consumer, .after = resource) 

igraph_psites <- graph_from_data_frame(d= igraph_links, # all links
                             vertices = nodes,
                             directed=T) # this makes sure resources (from) "point" to consumers (to)

## no parasites ##
igraph_links_np <- links_nopsites %>% 
  relocate(consumer, .after = resource)

igraph_nopsites <- graph_from_data_frame(d= igraph_links_np, # all links
                             vertices = nodes_nopsites,
                             directed=T) # this makes sure resources (from) "point" to consumers (to)

```

### I would first like to visualize degree (# links/node) distribution (PDF) and cumulative distribution (eCDF) for each web
```{r degree distribution}

# create degree and add it to nodes dfs:
deg.p <- igraph::degree(igraph_psites, mode = "all") 
nodes <- nodes %>% 
  mutate(degree = deg.p)

deg.np <- igraph::degree(igraph_nopsites, mode = "all") 
nodes_nopsites <- nodes_nopsites %>% 
  mutate(degree = deg.np)


# plot degree PDF for parasites web

ggplot(nodes, aes(x = degree)) +
  geom_histogram(binwidth = 1) + 
  geom_density()

# plot degree eCDF for parasites web
ggplot(nodes, aes(x = degree)) +
  geom_line(stat = "ecdf")+
  geom_point(stat = "ecdf")

# plot degree PDF for no parasites web
ggplot(nodes_nopsites, aes(x = degree)) +
  geom_histogram(binwidth = 1) + 
  geom_density()

# plot degree eCDF for no parasites web
ggplot(nodes_nopsites, aes(x = degree)) +
  geom_line(stat = "ecdf")+
  geom_point(stat = "ecdf")

# turns out package NetworkExtinction has a really nice function for this as well (oh oops so does igraph and cheddar)
# parasites #
psites_adj<- as.network(as_adjacency_matrix(igraph_psites), matrix.type = "adjacency")

NetworkExtinction::DegreeDistribution(psites_adj)

# without #
nopsites_adj<- as.network(as_adjacency_matrix(igraph_nopsites), matrix.type = "adjacency")
NetworkExtinction::DegreeDistribution(nopsites_adj)

```

## Ok, so for actually comparing webs, Dana compared the following:
- # nodes
- # links
- link density
- Connectance
- Adjusted connectance (Lafferty et al 2006)
- Longest chain
- Mean degree
- Sd degree
- Mean generality
- Sd generality
- Mean vulnerability
- Sd vulnerability

### Other metrics of interest:
  - mean centrality
(Williams and Martinez 2004)
	- Degree of caniballism
	- Degree of omnivory
	- Degree of looping
  - Degree of trophic similarity

### For parasite web
```{r calculating metrics parasites}

########## Parasite Web ########## 
# total potential links (m)
m <- vcount(igraph_psites)^2

# link density
link_dens.p <- edge_density(igraph_psites, loops = TRUE) # links/# nodes

# connectance # links/potential links
connectance <- vcount(igraph_psites)/m

# longest chain
longest.p <- diameter(igraph_psites)

# mean chain length:
mean_chain <- mean_distance(igraph_psites, directed = TRUE, details = TRUE)
average_chain_length_p <- mean_chain$res
## cannot find a way to extract any measure of variance

# histogram of chain lengths 1-8
dtable.p <- distance_table(igraph_psites, directed = TRUE)

# matrix of all distances between all nodes
dist.p <- distances(igraph_psites, algorithm = "dijkstra")

# mean degree
mean_deg.p <- mean(deg.p)
sd_deg.p <- sd(deg.p)

# mean generality
out.p <- igraph::degree(igraph_psites, mode = "out")
mean_gen.p <- mean(out.p)
sd_gen.p <- sd(out.p)

# mean vulnerability
inn <- igraph::degree(igraph_psites, mode = "in")
mean_vul.p <- mean(inn)
sd_vul.p <- sd(inn)

```

### For free no parasites/living web
```{r calculating metrics free living}

########## Parasite Web ########## 
# total potential links (m)
m.np <- vcount(igraph_nopsites)^2

# link density
link_dens.np <- edge_density(igraph_nopsites, loops = TRUE) # links/# nodes

# connectance # links/potential links
connectance.np <- vcount(igraph_nopsites)/m.np

# longest chain
longest.np <- diameter(igraph_nopsites)

# mean chain length:
mean_chain.np <- mean_distance(igraph_nopsites, directed = TRUE, details = TRUE)
average_chain_length.np <- mean_chain.np$res
## cannot find a way to extract any measure of variance

# histogram of chain lengths 1-8
dtable.np <- distance_table(igraph_nopsites, directed = TRUE)

# matrix of all distances between all nodes
dist.np <- distances(igraph_nopsites, algorithm = "dijkstra")

# mean degree
mean_deg.np <- mean(deg.np)
sd_deg.np <- sd(deg.np)

# mean generality
out.np <- igraph::degree(igraph_nopsites, mode = "out")
mean_gen.np <- mean(out.np)
sd_gen.np <-sd(out.np)

# mean vulnerability
in.np <- igraph::degree(igraph_nopsites, mode = "in")
mean_vul.np <- mean(in.np)
sd_vul.np <- sd(in.np)
```



### Make a nice table comparing values
```{r psite vs free living table}

values <- c("Total Possible Links", "Link Density", "Connectance", "Longest Chain", "Mean Chain Length", "Mean Degree", "SD Degree","Mean Generality", "SD Generality", "Mean Vulnerability", "SD Vulnerability")
parasite <- c(m, link_dens.p, connectance, longest.p, average_chain_length_p, mean_deg.p, sd_deg.p, mean_gen.p, sd_gen.p, mean_vul.p, sd_vul.p)
free_living <- c(m.np, link_dens.np, connectance.np, longest.np, average_chain_length.np, mean_deg.np, sd_deg.np, mean_gen.np, sd_gen.np, mean_vul.np, sd_vul.np)

psite_comp_table <- tibble(values, parasite, free_living)
kable(psite_comp_table) %>% 
  kable_classic(html_font = "Cambria")
```
## Now compare North vs South
### First make subweb data
```{r make no parasite subwebs}
######################################
###########     SOUTH    #############
######################################

########### filter north out to create new node list ########### 
nodes_south <- nodes %>% 
  filter(province == "S"|
         province == "Central"|
         province == "Both")

###########  subset the web using this new node list ########### 
# we only want consumers/resource links in the web if those consumers/resources are in our clean node list
# we can do this in five steps

#1. create a "keep" list of consumers
consumers_keep_s <- nodes_south %>% 
  dplyr::select(node) %>%  # we only need the unique identifier, "node", for matching
  rename(consumer = node) # the unique identifier of consumers in the link list is "consumer", and these colnames must match

#2. create a version of the link list that only has links with consumers in our "keep" list, e.g. south organisms only
consumer_keep_links_s <- inner_join(consumers_keep_s, links, by = join_by(consumer)) # inner join keeps only rows in 'links' that match 'consumers_keep'
  
#3. create a "keep" list of resources (this is identical to our consumer keep list, we just need to change the column name)
resources_keep_s <- consumers_keep_s %>% 
  rename(resource = consumer) # the unique identifier of resources in the link list is "resource", and these colnames must match

#4. use 'inner_join()' to create a link list from the consumer_keep_links list that ALSO only has resources from the keep list, e.g. south organisms
links_south <- inner_join(resources_keep_s, consumer_keep_links_s, by = join_by(resource))

######################################
###########     NORTH    #############
######################################

nodes_north <- nodes %>% 
  filter(province == "N"|
         province == "Central"|
         province == "Both")

###########  subset the web using this new node list ########### 
# we only want consumers/resource links in the web if those consumers/resources are in our clean node list
# we can do this in five steps

#1. create a "keep" list of consumers
consumers_keep_n <- nodes_north %>% 
  dplyr::select(node) %>%  # we only need the unique identifier, "node", for matching
  rename(consumer = node) # the unique identifier of consumers in the link list is "consumer", and these colnames must match

#2. create a version of the link list that only has links with consumers in our "keep" list, e.g. free living organisms only
consumer_keep_links_n <- inner_join(consumers_keep_n, links, by = join_by(consumer)) # inner join keeps only rows in 'links' that match 'consumers_keep'
  
#3. create a "keep" list of resources (this is identical to our consumer keep list, we just need to change the column name)
resources_keep_n <- consumers_keep_n %>% 
  rename(resource = consumer) # the unique identifier of resources in the link list is "resource", and these colnames must match

#4. use 'inner_join()' to create a link list from the consumer_keep_links list that ALSO only has resources from the keep list, e.g. free living organisms
links_north <- inner_join(resources_keep_n, consumer_keep_links_n, by = join_by(resource))
```


```{r}

```


```{r graveyard}

##### attempts to use cheddar for chain length calculations failed because of web size #####

#options(cheddarMaxQueue=0)

# We remove cannibalistic links to avoid entering an infinte loop when calculating path lengths
#ca_riz_parasites1 <- RemoveCannibalisticLinks(ca_riz_parasites, title='no cannibals')
#ca_riz_parasites_ag <- LumpTrophicSpecies(ca_riz_parasites, include.isolated = FALSE, title = 'trophic sp', weight.by = NULL)

#QuantitativeDescriptors(ca_riz_parasites_ag, weight = 'confidence')
#TrophicChainsStats(ca_riz_parasites_ag)


```

















