---
title: "California Rocky Intertidal Food Web"
output: html_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(igraph) 
library(network) 
library(sna)
library(ggraph)
library(visNetwork)
library(threejs)
library(networkD3)
library(ndtv) 
library(tidyverse)
library(readxl)
library(NetIndices)
library(visNetwork)

```

# Welcome to the California Rocky Intertidal Food Web interactive visualization tool!

### This food web network was created as part of an ongoing research project based in the Parasite Ecology Lab at the University of California Santa Barbara. We have compiled interaction reports from published scientific literature, intertidal guidebooks, our own field and lab studies, and expert testimony to generate this network. The network currently represents **544 species** and **over 1000 unique trophic links**, including grazing, predation, scavenging, and parasitism. We add data to the network on a daily basis!

#### To use this tool, simply select your species of interest using the drop down menu. You can also interact with the web by hovering over individual "nodes", which will display the associated species or group name. Clicking on nodes will highlight all links connected to that node. You can also drag the nodes around and change the shape of the network display!

#### *If you have questions about this ongoing research or wish to report a species interaction for inclusion in the web, please contact us by emailing our project lead, Zoe Zilz, at [zilz\@ucsb.edu](mailto:zilz@ucsb.edu){.email}*.

*This project has been generously funded by the Worster Family and the UCSB Coastal Fund*

```{r echo=FALSE, message=FALSE, warning=FALSE}
edges_unclean <- read_csv("data/edges.csv")

edges_unclean$resourceNum <- as.numeric(edges_unclean$resourceNum) 
edges_unclean$consumerNum <- as.numeric(edges_unclean$consumerNum) 

cnodes <- edges_unclean %>% 
  dplyr::select(consumerNum, consumerName, consumerStage, consumerCat) %>% 
  rename_with(~str_remove(., 'consumer')) 

rnodes <- edges_unclean %>% 
  dplyr::select(resourceNum, resourceName, resourceStage, resourceCat) %>% 
  rename_with(~str_remove(., 'resource')) 

pre_nodes <- bind_rows(cnodes, rnodes) %>% 
  filter(!is.na(Num)) %>%  
  filter(duplicated(Num) == FALSE) %>% 
  rename(id = Num)

node_attributes <- read_excel("/Users/zoe/Library/CloudStorage/GoogleDrive-zilz@ucsb.edu/My Drive/SCHOOL/PHD_AT_UCSB/RIZ Food Web/Data/Food Web Data/DO NOT EDIT/RIZspeciesList_MASTER.xlsx", sheet = "Both") %>% 
  rename(Name = species,
         range = 'N/S/Both', 
         trophic_strategy = 'trophic strategy')

#setting up a bunch of functional groups bc this is easier
#filter feeders#
ffs <- c("tunicate", "sponge", "hydroid", "barnacle", "mussel", "bivalve", "scallop", "bryozoan", "zooplankton", "hydrocoral", "clam", "entoproct", "boring clam", "vermetid", "soft coral", "coral", "sipunculid")

#predators#
preds <- c("sea star", "asteroid", "octopus", "nudibranch", "murex", "bird", "shark", "lobster", "whelk", "sea spider", "human", "mammal", "snail")

#primary producers#
pps <- c("algae", "phytoplankton", "corraline algae", "angiosperm", "dead stuff", "dinoflagellate", "gametes")

#herbivores#
herbs <- c("urchin", "topsnail", "limpet", "littorine", "slipper snail", "true limpet", "chiton", "sea hare", "microsnail", "isopod", "abalone")

#opportunistic omnivores#
oos <- c("anemone", "shrimp", "ghost shrimp", "scale worm", "true crab", "porcelain crab", "tanaid", "copepod", "other arthropod", "cucumber", "hermit crab", "anomuran", "flatworm", "turbellarian", "polychaete", "fish", "amphipod", "microorganism", "brittle star")

#all parasites#
paras <- c("parasitic copepod", "gregarine", "parasitic barnacle", "myxozoan", "coccidian", "trematode", "cestode", "fungus", "monogene", "ciliate", "leech", "sporozoan", "orthonectid", "rhombozoan", "nematode", "nemertean")

ded <- c("dead stuff")

nodes <- left_join(pre_nodes, node_attributes) %>%  
  filter(duplicated(id) == FALSE) %>%  
  dplyr::select(id, Name, category, range, Zone, trophic_strategy) %>% 
  mutate(trophic_strategy2 = case_when(
    category %in% ffs ~ "filter feeder",
    category %in% preds ~ "predator",
    category %in% pps ~ "primary producer",
    category %in% herbs ~ "herbivore",
    category %in% oos ~ "opportunistic omnivores",
    category %in% paras ~ "parasites",
    category %in% ded ~ "dead stuff"
  )) %>% 
  mutate(trophic_strategy3 = coalesce(trophic_strategy, trophic_strategy2)) %>% 
  mutate(parasitic  = if_else(trophic_strategy2 == "parasites", "y", "n")) %>% 
  filter(duplicated(id) == FALSE) 

edges_unclean -> links_unclean 


links <- links_unclean %>% 
  filter(!is.na(resourceNum)) %>% 
  filter(!is.na(consumerNum)) %>% 
  mutate(incl.code = replace_na(incl.code, 1)) %>% 
  filter(incl.code != 0, 
         incl.code != 5, 
         ) %>% 
  filter(interaction.type != 2, #epibiont
         interaction.type !=9, #endocommensal
         interaction.type !=10, #ectocommensal
         interaction.type !=22, #mutualism
         interaction.type !=28) %>%  #boring
  dplyr::select(resourceNum,consumerNum,  interaction.type) %>%  
  mutate(parasitic = case_when(
    interaction.type > 4 & interaction.type < 8 ~ "y",
    interaction.type == 12 ~ "y",
    interaction.type > 23 & interaction.type < 26 ~ "y",
    TRUE ~ "n"
  ))

net <- graph_from_data_frame(d=links, vertices=nodes, directed=T)
deg <- igraph::degree(net, mode = "all") 

RIZadjmatrix <- as_adjacency_matrix(net, sparse = FALSE) 
trophRIZ<-TrophInd(RIZadjmatrix)
trophic_level <- trophRIZ$TL

vis.links <- links %>% 
  rename(to = consumerNum) %>% 
  rename(from = resourceNum) 

vis.nodes <- nodes %>% 
  mutate(degree = deg) %>%  
  mutate(trophic = trophic_level) %>% 
  mutate(color = case_when(
    trophic_strategy2 == "parasites" ~ "red",
    category == "algae" ~ "lightgreen",
    category == "phytoplankton" ~ "lightgreen",
    category == "angiosperms" ~ "lightgreen",
    category == "corraline algae" ~ "lightgreen",
    category == "dinoflagellate" ~ "lightgreen",
    category == "dead stuff" ~ "tan"
  ))%>% 
  mutate(level = trophic_level) # added this in on 2.22.23 to see if it could translate to a trophic level web in VisNetwork via HierarchicalLayout

vis.nodes$shape  <- "dot"  
vis.nodes$shadow <- TRUE # Nodes will drop shadow
vis.nodes$title  <- vis.nodes$Name # Text on click
vis.nodes$size   <- vis.nodes$degree*3 # Node size 
vis.nodes$borderWidth <- 2

visNetwork(vis.nodes, vis.links, height = "700px", width = "100%") %>% 
  visHierarchicalLayout(levelSeparation = 1000) %>% 
  #visIgraphLayout(layout = "layout.circle") %>% 
    visOptions(selectedBy =  "Name",
             highlightNearest = list(enabled = T, degree = 0)
             #nodesIdSelection = TRUE
             ) %>%
  visPhysics(enabled = FALSE)
```
