---
title: "Sanak Foodweb Incorporation"
output: html_notebook
---
```{r packages}

library(tidyverse)
library(needs)
prioritize(dplyr) # prevents other packages from masking functions in dplyr (mostly select)

```

#Sanak Archipelago Food Web from:
### Dunne, J. A., Maschner, H., Betts, M. W., Huntly, N., Russell, R., Williams, R. J., & Wood, S. A. (2016). The roles and impacts of human hunter-gatherers in North Pacific marine food webs. Scientific Reports, 6(1), Article 1. https://doi.org/10.1038/srep21179

#### Permissions:
This work is licensed under a Creative Commons Attribution 4.0 International License. The images or other third party material in this article are included in the article’s Creative Commons license, unless indicated otherwise in the credit line; if the material is not included under the Creative Commons license, users will need to obtain permission from the license holder to reproduce the material. To view a copy of this license, visit http://creativecommons.org/licenses/by/4.0/

### As of 8 Oct 2024, decided to run this again and exclude nodes matched based on family. I will manually assess those nodes and determine whether the links between them should be included AS RESOURCES ONLY (e.g. a specific bird species eats Staphylinidae)
### I am also using the most updated node list
```{r import sanak data and my data}

## my node list
#ca_nodes <- read_csv("data/archives/RIZspecieslist_fulltaxonomy_28may24.csv") %>% < original
ca_nodes <- read_csv("data/RIZspecieslist_fulltaxonomy_25sep24.csv") %>% # newe
  rename(Name = species,
         ca_id = id.x) # so it matches name

# sanak node list 

sanak_nodes_all <- read_csv("data/sanak_web/sanak_nodelist.csv") %>% 
  rename(sanak_id = ID)

sanak_nodes_iz <- sanak_nodes_all %>% 
  filter(habitat == "intertidal")

# sanak iz web
sanak_iz <- read_csv("data/sanak_web/sanak_iz_links.csv")

```
## Step 1: Wrangle sanak nodes so they match my nodes
```{r node wrangling}

# first start by keeping species in my list that are present in sanak list
matched_species <- inner_join(sanak_nodes_iz, ca_nodes, by = join_by(Name)) %>% 
  mutate(matched_using = "species")

# now we will make a list of genera in the sanak list that are in my list 
sanak_genera <- sanak_nodes_iz %>% 
  rename(genus = Name) # so we can join on genus column in ca_nodes

matched_genera <- inner_join(sanak_genera, ca_nodes, by = join_by(genus))%>% 
  mutate(matched_using = "genus")


##### saving these to manually go through later #####
sanak_fams <- sanak_nodes_iz %>% 
  rename(family = Name) # so we can join on genus column in ca_nodes

matched_families <- inner_join(sanak_fams, ca_nodes, by = join_by(family))%>% 
  mutate(matched_using = "family")

#####################################################

# need to make a full list of all the matched taxa and include a column indicating how that taxon was matched (e.g. species, genus)

matched_nodes_all <- rbind(matched_species, matched_genera#, matched_families # no longer including these
                           )

matched_nodes_check <- matched_nodes_all %>% 
  dplyr::select(sanak_id, Name, matched_using)

```

## Step 2: Wrangling the sanak links to match my node info

### first i did this piecemeal for species:species first, then genus:genus, finally family:family, but then realized i need all interactions between matched taxa including species:species, species:genus, species: family, etc, so I went back and made a big combo list of all the taxa that matched from species:family level (i excluded higher levels based on logic). taxonomic level matches are below this next chunk

#### 361 nodes in the california web are represented in the sanak web, either by their exact species (n = 52), their genus (n = 111), or their family (n = 198). This leaves out nodes that might be the same but be called a different or old name. Oh well!

```{r link wrangling for matched genera and species}
# make a version of the matched nodes list to join with the link list (need to make sure IDs can bind)

# first for consumers
consumers_to_match <- matched_nodes_all %>% 
  rename(Consumer = sanak_id)  # ID column name in link list

consumers_data <- inner_join(sanak_iz, consumers_to_match, by = join_by(Consumer), relationship = "many-to-many") %>% 
  dplyr::select(Consumer, Name, ca_id, common_name, category, life_stage, Resource, matched_using) # only need some relevant data, can add more later

# then for resources
resources_to_match <- matched_nodes_all %>% 
  rename(Resource = sanak_id) %>%   # resource ID column name in link list, to match
  dplyr::select(Resource, Name, ca_id, common_name, category, life_stage, matched_using, zone)

# now put all links together
consumer_resource_links <- inner_join( # did left_join at first but it included a lot of resources that didn't match ca nodes
  consumers_data, resources_to_match, by = join_by(Resource), relationship = "many-to-many") # this matching should also automatically add .x to columns from consumers_data and .y to resources_to_match

```

## Step 3: Create links to add to main edge list
### MASTER CODE for manipulating dataframes to match with edgelist_master
```{r match edgelist_MASTER}

sanak_links_to_paste <- consumer_resource_links %>% 
  #filter(!is.na(resourceNum)) %>% ## we already did this by using inner join above
  
  #add in columns that are present in edgelist master:
  #mutate(date_added = "29 May 2024") %>% # original
  mutate(date_added = "8 Oct 2024") %>% 
  mutate(INITIALS = "ZZ_coded") %>% 
  
  # need to change the below so they are different based on what taxon they were matched using (species, genus, or family)
  mutate(justification = case_when(matched_using.x == "species" & matched_using.y == "species" ~ 3, # read in a paper
                                   .default = 6) # read in lit at higher taxonomic level
                                   ) %>% 
  mutate(incl.code = 2) %>% # both species in web but interaction reported elsewhere
  mutate(confidence = case_when(justification == 3 ~ 2,
                                justification == 6 ~ 3) # 2 both higher taxa in web but interaction reported elsewhere
           ) %>% 
  mutate(pools_exposed = NA) %>% # keep this empty for now
  mutate(localities = "Sanak Archipelago") %>%  
  mutate(MorrisHaderlie = "no") %>%
  mutate(epibiota_consumed = "y") %>% 
  mutate(notes = "") %>% 
  mutate(link.ID = (11849 + seq_len(length.out = n()))) %>%  # we deleted links # 11850 - 18984 from links_master on 8oct24 so this should just fill that gap
  mutate(ref_num = "375") %>% # ref nums for ONLY 375 = Dunne 2016 becuase that's where the data is from
  mutate(interaction.type = case_when(
    str_detect(category.y, "algae") ~ 3, 
    .default = 1)
  ) %>%  
  rename(consumerNum = ca_id.x, 
         consumerName = Name.x,
         consumerCat = category.x,
         resourceNum = ca_id.y,
         resourceName = Name.y,
         resourceCat = category.y
         ) %>% 
    
   # need to add a column that creates numeric life stage for consumers
  mutate(consumerStage = case_when(
    life_stage.x == "adult" | life_stage.x == "medusa" ~ 1,
    life_stage.x == "egg" ~ 2,
    life_stage.x == "juvenile" ~ 4, # juvenile fish/snail/crab
    life_stage.x == "hydroid" ~ 5, # polyp stage in cnidarians
    life_stage.x == "multiple" ~ 6, 
    .default = NA
  )) %>% 
  
  # need to add a column that creates numeric life stage for resources as well
  mutate(resourceStage = case_when(
    life_stage.y == "adult" | life_stage.y == "medusa" ~ 1,
    life_stage.y == "egg" ~ 2,
    life_stage.y == "juvenile" ~ 4, # juvenile fish/snail/crab
    life_stage.y == "small" ~ 4, # juvenile other
    life_stage.y == "nauplii" ~ 3, # first stage larva
    life_stage.y == "hydroid" | life_stage.y == "hydroid colony" ~ 5, # polyp stage in cnidarians
    life_stage.y == "multiple" ~ 6,
    .default = NA
  )) %>% 

  # now select out all the columns that don't match and put them in the right order
  select(INITIALS, consumerNum, consumerName, consumerCat, consumerStage, resourceNum, resourceName, resourceCat, resourceStage, interaction.type, confidence, justification, incl.code, ref_num, localities, zone, pools_exposed, MorrisHaderlie, epibiota_consumed, notes, link.ID, date_added)
                                                          

```

## Step 4: export to CSV
```{r create csv to merge with edgelist master}

write_csv(sanak_links_to_paste, "data/sanak_links_to_paste_8oct2024.csv")

```

## Step 5: Generate links from nodes matched by family to assess manually
basically repeat steps 3 - 4 for the matched_families dataframe
```{r link wrangling for matched families}
# make a version of the matched nodes list to join with the link list (need to make sure IDs can bind)

# first for consumers
consumers_to_match_fams <- matched_families %>% 
  rename(Consumer = sanak_id)  # ID column name in link list

consumers_data_fams <- inner_join(sanak_iz, consumers_to_match_fams, by = join_by(Consumer), relationship = "many-to-many") %>% 
  dplyr::select(family, Consumer, Name, ca_id, common_name, category, life_stage, Resource, matched_using) # only need some relevant data, can add more later

# then for resources
resources_to_match_fams <- matched_families %>% 
  rename(Resource = sanak_id) %>%   # resource ID column name in link list, to match
  dplyr::select(family, Resource, Name, ca_id, common_name, category, life_stage, matched_using, zone)

# now put all links together
consumer_resource_links_fam <- inner_join( # did left_join at first but it included a lot of resources that didn't match ca nodes
  consumers_data_fams, resources_to_match_fams, by = join_by(Resource), relationship = "many-to-many") # this matching should also automatically add .x to columns from consumers_data and .y to resources_to_match

```
### MASTER CODE for manipulating dataframes to match with edgelist_master
```{r match family nodes to edgelist_MASTER}

sanak_links_to_paste_fams <- consumer_resource_links_fam %>% 
  #filter(!is.na(resourceNum)) %>% ## we already did this by using inner join above
  
  #add in columns that are present in edgelist master:
  #mutate(date_added = "29 May 2024") %>% # original
  mutate(date_added = "8 Oct 2024") %>% 
  mutate(INITIALS = "ZZ_coded") %>% 
  
  # need to change the below so they are different based on what taxon they were matched using (species, genus, or family)
  mutate(justification = 6 ) %>% # read in lit at higher taxonomic level
  mutate(incl.code = 2) %>% # both species in web but interaction reported elsewhere
  mutate(confidence =  3) %>% # 2 both higher taxa in web but interaction reported elsewhere
  mutate(pools_exposed = NA) %>% # keep this empty for now
  mutate(localities = "Sanak Archipelago") %>%  
  mutate(MorrisHaderlie = "no") %>%
  mutate(epibiota_consumed = "y") %>% 
  mutate(notes = "") %>% 
  mutate(link.ID = (13723 + seq_len(length.out = n()))) %>%  # we deleted links # 11850 - 18984 from links_master, and last dataframe added back 11850-13742
  mutate(ref_num = "375") %>% # ref nums for ONLY 375 = Dunne 2016 becuase that's where the data is from
  mutate(interaction.type = case_when(
    str_detect(category.y, "algae") ~ 3, 
    .default = 1)
  ) %>%  
  rename(consumerNum = ca_id.x, 
         consumerName = Name.x,
         consumerCat = category.x,
         resourceNum = ca_id.y,
         resourceName = Name.y,
         resourceCat = category.y,
         consumerFam = family.x,
         resourceFam = family.y # keepign these in so we can assess later
         ) %>% 
    
   # need to add a column that creates numeric life stage for consumers
  mutate(consumerStage = case_when(
    life_stage.x == "adult" | life_stage.x == "medusa" ~ 1,
    life_stage.x == "egg" ~ 2,
    life_stage.x == "juvenile" ~ 4, # juvenile fish/snail/crab
    life_stage.x == "hydroid" ~ 5, # polyp stage in cnidarians
    life_stage.x == "multiple" ~ 6, 
    .default = NA
  )) %>% 
  
  # need to add a column that creates numeric life stage for resources as well
  mutate(resourceStage = case_when(
    life_stage.y == "adult" | life_stage.y == "medusa" ~ 1,
    life_stage.y == "egg" ~ 2,
    life_stage.y == "juvenile" ~ 4, # juvenile fish/snail/crab
    life_stage.y == "small" ~ 4, # juvenile other
    life_stage.y == "nauplii" ~ 3, # first stage larva
    life_stage.y == "hydroid" | life_stage.y == "hydroid colony" ~ 5, # polyp stage in cnidarians
    life_stage.y == "multiple" ~ 6,
    .default = NA
  )) %>% 

  # now select out all the columns that don't match and put them in the right order
  select(INITIALS, consumerFam, consumerNum, consumerName, consumerCat, consumerStage, resourceFam, resourceNum, resourceName, resourceCat, resourceStage, interaction.type, confidence, justification, incl.code, ref_num, localities, zone, pools_exposed, MorrisHaderlie, epibiota_consumed, notes, link.ID, date_added)
                                                          
######### FAMILY only CSV ######### 
write_csv(sanak_links_to_paste_fams, "data/sanak_fams_to_review_8oct2024.csv")

```


### deprecated code since i combined all matched nodes into one dataframe

```{r link wrangling for matched species, include= FALSE}
# make a version of the matched species list to join with the link list (need to make sure IDs can bind)

# first for consumers
consumers_to_match_s <- matched_species %>% 
  rename(Consumer = sanak_id)  # ID column name in link list

consumers_data_s <- inner_join(sanak_iz, consumers_to_match_s, by = join_by(Consumer), relationship = "many-to-many") %>% 
  select(Consumer, Name, ca_id, common_name, category, life_stage, Resource) # only need some relevant data, can add more later

# then for resources
resources_to_match_s <- matched_species %>% 
  rename(Resource = sanak_id) %>%   # resource ID column name in link list, to match
  select(Resource, Name, ca_id, common_name, category, life_stage, zone)


consumers_resources_s <- inner_join( # did left_join at first but it included a lot of resources that didn't match ca nodes
  consumers_data_s, resources_to_match_s, by = join_by(Resource), relationship = "many-to-many") # this matching should also automatically add .x to columns from consumers_data and .y to resources_to_match

```

```{r link wrangling for matched genera, include= FALSE}
# make a version of the matched gen list to join with the link list (need to make sure IDs can bind)

# first for consumers
consumers_to_match_g <- matched_genera %>% 
  rename(Consumer = sanak_id)  # ID column name in link list

consumers_data_g <- inner_join(sanak_iz, consumers_to_match_g, by = join_by(Consumer), relationship = "many-to-many") %>% 
  select(Consumer, Name, ca_id, common_name, category, life_stage, Resource) # only need some relevant data, can add more later

# then for resources
resources_to_match_g <- matched_genera %>% 
  rename(Resource = sanak_id) %>%   # resource ID column name in link list, to match
  select(Resource, Name, ca_id, common_name, category, life_stage, zone)


consumers_resources_g <- inner_join( # did left_join at first but it included a lot of resources that didn't match ca nodes
  consumers_data_g, resources_to_match_g, by = join_by(Resource), relationship = "many-to-many") # this matching should also automatically add .x to columns from consumers_data and .y to resources_to_match

```

```{r match edgelist_MASTER species, include= FALSE}

sanak_links_to_paste_species <- consumer_resources %>% 
  #filter(!is.na(resourceNum)) %>% ## we already did this by using inner join above
  
  #add in columns that are present in edgelist master:
  mutate(date_added = "29 May 2024") %>% 
  mutate(INITIALS = "ZZ_coded") %>% 
  mutate(justification = 3) %>% # read in a paper
  mutate(confidence = 2) %>% # both species in web but interaction reported elsewhere
  mutate(pools_exposed = NA) %>% # keep this empty for now
  mutate(incl.code = 2) %>% # both species in web but interaction reported elsewhere
  mutate(localities = "Sanak Archipelago") %>%  
  mutate(MorrisHaderlie = "no") %>%
  mutate(epibiota_consumed = "y") %>% 
  mutate(notes = "") %>% 
  mutate(link.ID = (11849 + seq_len(length.out = n()))) %>%  # 11849 is currently the last link id # in edgelist_master so make sure to change
  mutate(ref_num = 375) %>% # ref num for sanak web paper
  mutate(interaction.type = case_when(
    str_detect(category.y, "algae") ~ 3, 
    .default = 1
  )
  ) %>%  
  rename(consumerNum = ca_id.x, 
         consumerName = Name.x,
         consumerCat = category.x,
         resourceNum = ca_id.y,
         resourceName = Name.y,
         resourceCat = category.y
         ) %>% 
    
   # need to add a column that creates numeric life stage for consumers
  mutate(consumerStage = case_when(
    life_stage.x == "adult" | life_stage.x == "medusa" ~ 1,
    life_stage.x == "egg" ~ 2,
    life_stage.x == "juvenile" ~ 4, # juvenile fish/snail/crab
    life_stage.x == "hydroid" ~ 5, # polyp stage in cnidarians
    life_stage.x == "multiple" ~ 6, 
    .default = NA
  )) %>% 
  
  # need to add a column that creates numeric life stage for resources as well
  mutate(resourceStage = case_when(
    life_stage.y == "adult" | life_stage.y == "medusa" ~ 1,
    life_stage.y == "egg" ~ 2,
    life_stage.y == "juvenile" ~ 4, # juvenile fish/snail/crab
    life_stage.y == "small" ~ 4, # juvenile other
    life_stage.y == "nauplii" ~ 3, # first stage larva
    life_stage.y == "hydroid" | life_stage.y == "hydroid colony" ~ 5, # polyp stage in cnidarians
    life_stage.y == "multiple" ~ 6,
    .default = NA
  )) %>% 

  # now select out all the columns that don't match and put them in the right order
  select(INITIALS, consumerNum, consumerName, consumerCat, consumerStage, resourceNum, resourceName, resourceCat, resourceStage, interaction.type, confidence, justification, incl.code, ref_num, localities, zone, pools_exposed, MorrisHaderlie, epibiota_consumed, notes, link.ID, date_added)
                                                          

```

```{r match edgelist_MASTER genera, include= FALSE}

sanak_links_to_paste_genera <- consumers_resources_g %>% 
  #filter(!is.na(resourceNum)) %>% ## we already did this by using inner join above
  
  #add in columns that are present in edgelist master:
  mutate(date_added = "29 May 2024") %>% 
  mutate(INITIALS = "ZZ_coded") %>% 
  mutate(justification = 6) %>% # more broadly in literature (e.g. genera)
  mutate(confidence = 3) %>% # both broader taxa in web but interaction reported elsewhere
  mutate(pools_exposed = NA) %>% # keep this empty for now
  mutate(incl.code = 2) %>% # both species in web but interaction reported elsewhere
  mutate(localities = "Sanak Archipelago") %>%  
  mutate(MorrisHaderlie = "no") %>%
  mutate(epibiota_consumed = "y") %>% 
  mutate(notes = "") %>% 
  mutate(link.ID = (12012 + seq_len(length.out = n()))) %>%  # added to end of _species 
  mutate(ref_num = 375) %>% # ref num for sanak web paper
  mutate(interaction.type = case_when(
    str_detect(category.y, "algae") ~ 3, 
    .default = 1
  )
  ) %>%  
  rename(consumerNum = ca_id.x, 
         consumerName = Name.x,
         consumerCat = category.x,
         resourceNum = ca_id.y,
         resourceName = Name.y,
         resourceCat = category.y
         ) %>% 
    
   # need to add a column that creates numeric life stage for consumers
  mutate(consumerStage = case_when(
    life_stage.x == "adult" | life_stage.x == "medusa" ~ 1,
    life_stage.x == "egg" ~ 2,
    life_stage.x == "juvenile" ~ 4, # juvenile fish/snail/crab
    life_stage.x == "hydroid" ~ 5, # polyp stage in cnidarians
    life_stage.x == "multiple" ~ 6, 
    .default = NA
  )) %>% 
  
  # need to add a column that creates numeric life stage for resources as well
  mutate(resourceStage = case_when(
    life_stage.y == "adult" | life_stage.y == "medusa" ~ 1,
    life_stage.y == "egg" ~ 2,
    life_stage.y == "juvenile" ~ 4, # juvenile fish/snail/crab
    life_stage.y == "small" ~ 4, # juvenile other
    life_stage.y == "nauplii" ~ 3, # first stage larva
    life_stage.y == "hydroid" | life_stage.y == "hydroid colony" ~ 5, # polyp stage in cnidarians
    life_stage.y == "multiple" ~ 6,
    .default = NA
  )) %>% 

  # now select out all the columns that don't match and put them in the right order
  select(INITIALS, consumerNum, consumerName, consumerCat, consumerStage, resourceNum, resourceName, resourceCat, resourceStage, interaction.type, confidence, justification, incl.code, ref_num, localities, zone, pools_exposed, MorrisHaderlie, epibiota_consumed, notes, link.ID, date_added)

```